Warming Degrees and Flowering Trees

### 2025-04-04 paper/paper.qmd
finalest final commit

The regression analysis in @tbl-reg-results reveals a significant inverse relationship between March temperature and cherry blossom blooming dates. For each 1°C increase in March temperature, cherry blossoms bloom approximately 2.25 days earlier (p < 0.01). This relationship remains consistent across both standard and robust OLS models, though the latter shows larger standard errors, suggesting heteroscedasticity in the data. With an R-squared of 0.32, March temperature appears to explain about one-third of the variation in blooming dates. When March temperature is 0°C, blossoms are predicted to emerge on the 88th day of the year (late March). These findings indicate that spring temperature serves as a meaningful predictor of cherry blossom phenology. Diagnostic plots are in @sec-appendix-1

@@

# Simple scatterplot with regression line
plt.figure(figsize=(8, 4))
plt.scatter(df['march_temp_c'], df['flower_doy'], color='#FF9ECD', edgecolor='#E75480', alpha=0.7)
plt.plot(sorted(df['march_temp_c']), model_0.predict(sm.add_constant(sorted(df['march_temp_c']))), color='red')
plt.text(0.1, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)
plt.title("Sakura Blooming Day vs March Temperature")
plt.xlabel("March Avg Temperature (°C)")
plt.ylabel("Blooming Day of Year (DoY)")
plt.show()
```

```{python}
#| label: fig-train-plt
#| fig-cap: "Trained Predictions Visualization"
#| message: false
#| echo: false
#| warning: false
#| fig-align: center

# Test OLS model (nonrobust)

model_test = model_0

# Prepare test data
X_test = sm.add_constant(test_df['march_temp_c'])
y_test = test_df['flower_doy']

# Predict on test data
y_pred = model_test.predict(X_test)

# Plot actual vs predicted
plt.figure(figsize=(10, 6))
plt.plot(test_df['year'], y_test, label='Actual Blooming DoY', marker='o')
plt.plot(test_df['year'], y_pred, label='Predicted Blooming DoY', marker='x')
plt.xlabel("Year")
plt.ylabel("Blooming Day of Year")
plt.title("Actual vs Predicted Sakura Blooming (2001–2019)")
plt.legend()
plt.grid(True)
```

@@

>>>>

The regression analysis in @tbl-reg-results, visualized in @fig-reg-plt, reveals a significant inverse relationship between March temperature and cherry blossom blooming dates. For each 1°C increase in March temperature, cherry blossoms bloom approximately 2.25 days earlier (p < 0.01). This relationship remains consistent across both standard and robust OLS models, though the latter shows larger standard errors, suggesting heteroscedasticity in the data. With an R-squared of 0.32, March temperature appears to explain about one-third of the variation in blooming dates. When March temperature is 0°C, blossoms are predicted to emerge on the 88th day of the year (late March). These findings indicate that spring temperature serves as a meaningful predictor of cherry blossom phenology. Diagnostic plots are in @sec-appendix-1

@@

# Simple scatterplot with regression line
plt.figure(figsize=(8, 4))
plt.scatter(df['march_temp_c'], df['flower_doy'], color='#FF9ECD', edgecolor='#E75480', alpha=0.7)

plt.plot(sorted(df['march_temp_c']), model_0.predict(sm.add_constant(sorted(df['march_temp_c']))), color='red')

plt.text(0.1, 95, f"blooming_day = {model_0.params[1]:.2f} * temp + {model_0.params[0]:.2f}", color="red", fontsize=12)

plt.title("Sakura Blooming Day vs March Temperature")
plt.xlabel("March Avg Temperature (°C)")
plt.ylabel("Blooming Day of Year (DoY)")
plt.show()
```

@@

## Code for Plots

The code used to render the plots and tables in this paper can be found at <https://github.com/lcarnegie/sakura/tree/main/scripts>
